Supplement to: Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-09-22 supplement.tex
really terrible overview of fragment grammars

\subsection{Structure Learning}

We use Alg.~\ref{grammarInductionAlgorithm} to search for the structure of the DSL that best explains the frontiers.

\begin{figure}[H]
  \centering
  \begin{minipage}{0.675\textwidth}    
    \begin{algorithm}[H]
    \setcounter{algorithm}{2} % there are two algorithms in the main paper.
       \caption{DSL Induction Algorithm}
       \label{grammarInductionAlgorithm}
       \begin{algorithmic}
         \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
         \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
         \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[p|\mathcal{D},\theta]$
         \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
         \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - \|\theta\|_0$
         \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
         \WHILE {true}
         \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, p\in \mathcal{F}_x, s\text{ a fragment  of }p\}$
         \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
         \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
         \STATE $\mathcal{D}\gets\mathcal{D}'$
         \ENDWHILE
       \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\end{figure}

>>>>

\subsection{Fragment grammars}

`Fragment grammars'~\cite{tim} is a formalism from computational linguistics
developed for the purpose of modeling
the reuse of structure
in natural language.
Here,
we will adapt the fragment grammar formalism
to model the reuse of structure in
a formal language ($\lambda$ calculus),
interpreting reuse of structure as subroutines.

In the original fragment grammar formalism,
a fragment grammar
contains a probabilistic CFG called its \emph{base grammar};
a \emph{fragment} is a tree drawn from the base grammar,
but which can contain nonterminal symbols.
As a motivating example,
consider the following base grammar:
\begin{align*}
  S\to&(+ \;S\; S)\\
  S\to&(\times \;S  \;S)\\
  S\to&0\\
      S\to&1  
\end{align*}
An example expression drawn from this grammar is $(+ \;1\; (\times \;0 \;0))$.
An example \emph{fragment} drawn from this grammar $(+ \;1\; (\times \;S\; S))$.

When extending fragment grammars to $\lambda$-calculus we write nonterminal symbols as free variables,
and use the (current) DSL as our base grammar.
For example, if our DSL includes addition, multiplication, and the numbers zero \& one (as in our example base grammar),
then a possible  $\lambda$-fragment would be $(+ \;1\; (\times \;x\; y))$.

In order to use fragments
for DSL induction,
we need several
pieces: (1) a procedure for proposing fragments from programs found in the frontiers (described in figure~\ref{extractFragments});
(2) a procedure for evaluating
the likelihood of a program given that a fragment
has been added to the DSL (defined in figure~\ref{match});
and (3) a procedure for converting a fragment to a subroutine (i.e., a closed, typed $\lambda$-calculus expression; described in Figure~\ref{close}).
Putting these procedures together, Algorithm~\ref{grammarInductionAlgorithm}
is used to induce a new DSL.

%\vfill\pagebreak
\begin{figure}  
  \begin{minipage}[c]{\textwidth}\toprule 
    \begin{align*}
  \text{fragments}(\lambda z.e)& = \text{fragments}'(\lambda z.e)\cup \text{fragments}(e)\\
  \text{fragments}(f \; x)& = \text{fragments}'(f \; x)\cup \text{fragments}(f)\cup\text{fragments}(x)\\
  \text{fragments}(e)& = \varnothing\text{, if $e$ is a variable or primitive.}\\
  \\
  \text{fragments}'(\lambda z.e)& = \left\{\lambda z.e' | e'\in \text{fragments}'(e) \right\}\cup \left\{v \right\}\text{, $v$ a free variable}\\
  \text{fragments}'(f\; x)& = \left\{f'\; x' | f'\in \text{fragments}'(f),\;x'\in \text{fragments}'(x) \right\}\cup \left\{v \right\}\text{, $v$ a free variable}\\
    \text{fragments}'(e)& = \left\{e\right\}\cup \left\{v \right\}\text{, $v$ a free variable, $e$ a variable or primitive}
\end{align*}\bottomrule 
    \end{minipage}
\caption{Procedure for extracting possible fragments from a program. To prevent an explosion in the number of fragments, we also cap the number of free variables in a fragment to 3.}\label{extractFragments}
\end{figure}
\begin{figure}
  \toprule \begin{minipage}[c]{\textwidth}
    \begin{align*}
      \text{closeFragment}(e)& = \text{closeFragment}(\lambda z.e)\text{, if $z$ is a free variable in $e$}\\
      \text{closeFragment}(e)& = e\text{, if $e$ has no free variables.}
    \end{align*}
    \bottomrule 
  \end{minipage}
  \caption{Procedure for converting a fragment to a subroutine.}\label{close}
\end{figure}

\begin{figure}
  \toprule \begin{minipage}[c]{\textwidth}
    \begin{align*}
      \probability[p|\mathcal{D},\theta]& = P_{\mathcal{D},\theta} (p|\tau,\varnothing )\text{, where $p:\tau$}\\
      P_{\mathcal{D},\theta} (\lambda z.p|\alpha\to\beta,\mathcal{E})& =
      P_{\mathcal{D},\theta}(p|\beta,[\tau\mapsto\alpha]\mathcal{E})\text{, where $\tau$ a fresh type variable}\\
      P_{\mathcal{D},\theta}(e|\tau,\mathcal{E})& = \underbrace{\sum_{f,xs\in \text{parse}(e)} P'_{\mathcal{D},\theta}(f,xs|\tau,\mathcal{E})}_{\text{Marginalizing out how $\mathcal{D}$ produced $e$}}\text{, when $e$ not a $\lambda$-abstraction}\\
      \\
      P'_{\mathcal{D},\theta}(f,xs|\tau,\mathcal{E})& = \underbrace{\sum_{f':\tau_1\to\tau_2\to\cdots\to\tau\in \mathcal{D}}\frac{\theta_f}{Z_{\tau,\mathcal{D},\theta,\mathcal{E}}}
        \indicator[\text{match}(f',f)\not=\bot]}_{\text{Marginalize over which fragment produced $f,xs$}}\times\\&\prod_{n=1}^{|xs|}P_{\mathcal{D},\theta}(x_n|\tau_n,\mathcal{E})\prod_{e,\tau'\in \text{match}(f',f)}P_{\mathcal{D},\theta}(e|\tau',\mathcal{E})\\
      \\
      \text{match}(\lambda z.b,\lambda z.b')& = \text{match}(b,b')\\
      \text{match}(f\;x,f'\;x')& = \text{match}(f,f')\cup\text{match}(x,x')\\
      \text{match}(p,p)& = \varnothing \text{, $p$ a primitive or variable bound in fragment}\\
      \text{match}(v,p)& = \left\{v\mapsto p   \right\} \text{, $v$ a variable free in fragment}\\
      \text{match}(e,e')& = \bot\text{, otherwise.}
      \\
      \\
      Z_{\tau,\mathcal{D},\theta,\mathcal{E}}& = \underbrace{\sum_{f:\tau_1\to\tau_2\to\cdots\to\tau\in \mathcal{D}\cup\mathcal{E}}\theta_f}_{\text{Normalizing constant for type $\tau$}}\\\\
      \text{parse}(f\;x)& = \left\{f',xs+[x] \;|\; f',xs\in \text{parse}(f) \right\}\cup\left\{f\;x \right\}\\
      \text{parse}(e)& = \left\{e,[] \right\} \text{, $e$ not an application}
    \end{align*}
  \end{minipage}
  \bottomrule
  \caption{Procedure for calculating likelihood of a program after a new fragment has been added to the DSL. For simplicity we omit typing context and reuse type variables to indicate unification, as is commonly done in e.g. Prolog.}\label{match}
  \end{figure}

\begin{figure}[H]
  \centering
  \begin{minipage}{\textwidth}    
    \begin{algorithm}[H]
    \setcounter{algorithm}{2} % there are two algorithms in the main paper.
       \caption{DSL Induction Algorithm}
       \label{grammarInductionAlgorithm}
       \begin{algorithmic}
         \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
         \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
         \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[p|\mathcal{D},\theta]$
         \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
         \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - \|\theta\|_0$
         \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
         \WHILE {true}
         \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, p\in \mathcal{F}_x, s\in \text{fragments}(p)\}$
         \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
         \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\left\{\text{closeFragment}(s)|s\in \mathcal{D} \right\},\theta^*(\mathcal{D})$
         \STATE $\mathcal{D}\gets\mathcal{D}'$
         \ENDWHILE
       \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\end{figure}
